---
title: 'Homework #2'
author: "Harsha Senapathi"
date: "10/3/2019"
output: github_document
---


# Loading Library
```{r}
library(tidyverse)
library(tidyr)
library(readxl)
```

#Problem 1

The import and manipulation of the Mr. Trash Wheel Dataset.

```{r p1_import_clean}
trashwheel = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                        sheet = 1, 
                        range = "A2:N408") %>% 
  janitor::clean_names() %>%
  select(-dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
  rename("weight" = weight_tons, "volume" = volume_cubic_yards) #renaming columns to more reasonable variable names
```

The columns weight_tons and volume_cubic_yards were renamed to weight and volume since mentioning units seems superfluous. It can be mentioned once in the data description to suffice for it.

Import and manipulation of the precipitation data from 2017 and 2018

## Precipitation Dataset

The following code trunk reads and cleans precipitation data for 2017 and 2018. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable.

```{r p1_data_import_precip}

data_prcpttn_2017 = 
    read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = '2017 Precipitation', range = "A2:B14") %>%
    janitor::clean_names() %>% 
    drop_na(total) %>% #delete rows without precipitation data
    mutate(year = 2017)
    
data_prcpttn_2018 = 
    read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = '2018 Precipitation', range = "A2:B14") %>%
    janitor::clean_names() %>% 
    drop_na(total) %>% #delete rows without precipitation data
    mutate(year = 2018)

# Combine datasets for years 2017 and 2018

data_prcpttn = bind_rows(data_prcpttn_2017, data_prcpttn_2018) %>% 
    mutate(month = month.name[month]) %>% 
    select(year, month, total)

# Check the combined dataset
head(data_prcpttn)
tail(data_prcpttn)

```

## Interpretation
There are `r nrow(data_trashwheel)` observations and `r ncol(data_trashwheel)` variables in the Mr. Trash Wheel dataset, and `r nrow(data_prcpttn)` observations and `r ncol(data_prcpttn)` variables in the precipitation dataset. Key variables in the Mr. Trash Wheel dataset include `weight_tons`, number of different items (such as `sports_balls`) and `homes_powered`. For the precipitation dataset, the key variable is `total`, which denotes the volume of precipitation for each month. The total precipitation in 2017 was `r sum(filter(data_prcpttn_2017, year == 2017)$total)`. The median number of sports balls in a dumpster in 2018 was `r median(filter(data_trashwheel, year == 2018)$sports_balls)`.

#Problem 2

```{r p2_import_clean}

#cleaning polls dataset
polls = read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv")%>%
  separate(mon, c("year", "month", "day"), 
           convert = TRUE)%>%
  mutate (month = month.name[month])%>%
  mutate(prez_dem = recode (prez_dem, '1' = "dem")) %>% 
  mutate(prez_gop = recode (prez_gop, '1' = "gop", '2' = "gop")) %>% 
  pivot_longer(
    cols = c("prez_dem", "prez_gop"), 
    values_to = "president", 
    values_drop_na = TRUE
  ) %>% 
  select(-day, -name)

#cleaning snp dataset
snp = read_csv(file = "./data/fivethirtyeight_datasets/snp.csv")%>%
    separate(date, c("year", "month", "day"), 
             convert = TRUE)%>%
   mutate (month = month.name[month])%>%
  select(year, month, -day, everything())

#cleaning and transforming unemployment data
unemployment = read_csv("./Data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(Jan:Dec, 
    names_to = "month",
    values_to = "rate"
  ) %>% 
  mutate(month = match(month, month.abb)) %>%
  mutate(month = month.name[month]) %>%
  mutate(Year = as.integer(Year)) %>%
  janitor::clean_names()

#combining data sets, to create a dataset with polls, snp and unemployent rate.
Final = polls%>%
  left_join( snp,
             by = c("year", "month"))
  left_join( unemployment,
             by = c("year", "month"))

```



